---
title: "Simulating scenarios: comparing log-distance to the nearest feature to the cumulative influence measure"
author: "Bernardo Niebuhr, Bram van Moorter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=2cm
output: 
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
library(NinaR)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/",
  fig.align="center",
  message=FALSE, 
  warning=FALSE
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())
```

## Introduction

There are a few big questions that underlie the assessment of the effects of anthropogenic infrastructure on wildlife.
When performing environmental impact assessments, one aims not only to to find which factors affect wildlife and how strongly, 
but also (i) at which spatial scale there are effects and (ii) how these effects sum and interact when (as it is often the case) there are multiple infrastructures and vectors of landscape modification. The first question is also knows as the Zone of Influence (ZoI) problem. The second is generally tackled in the context of cumulative impact assessment.

In the main text of this manuscript, we argue that the cumulative influence measure (CumInf) - proportional to the density of infrastructure features - might better represent the cumulative effect of multiple features in the landscape than considering only the distance from the feature nearest to a given location. This, however, might vary depending on the number of features in the landscape, how these features are distributed in space, and what is the ZoI of each of those features.

Here we simulate some landscapes with point-type infrastructure spread following different patterns, calculate the distance to the nearest feature and the CumInf of the features, at multiple scales/ZoI, and assess when and how these variables might represent different sources of spatial variation.

## Simulate landscapes

First we simulate some landscape using point-type infrastructure as an example. They could represent the spatial location of houses, cabins, or wind turbines, for example. To do this we'll use a few functions designed within the R package `oneimpact`. We also load spatial packages it depends on and other packages for data manipulation and plotting.


```{r load_packages, message=FALSE, warning=FALSE}
#Load packages

# data manipulation
library(tibble)
library(dplyr)
library(purrr)
library(ggplot2) # for plot
library(ggpubr) # for plot
library(lsr) # for correlations

# spatial packages
library(mobsim)
library(raster)
library(landscapetools)
library(sf)
library(spatstat)

# oneimpact package
library(oneimpact)
```

We set a 30x30 km^2^ landscape and can simulate points following different spatial patterns. Here is an example landscape where the points are spread close to a single center or cluster (e.g. houses in a village).

```{r one_point_pattern, fig.dim=c(5,5)}
set.seed(1234)

# simulate a single patch
nfeat <- 1000 # number of features
ext <- 30000 # extension of the landscape
nc <- 1 # number of centers or patches
wd <- ext/20 # width of the patch
pts <- set_points(n_features = 1000, centers = nc,
                  width = wd, res = 100,
                  extent_x = c(0, ext), extent_y = c(0, ext))

landscapetools::show_landscape(pts$rast, legend.position = "none")
```

We can now simulate landscapes with different patterns. We start with 3 spatial distribution of points: regular, random, and clumped distribution of points. For the clumped distribution, we use two scenarios, where the point features are spread in either 5 or only 1 focal cluster, and we choose two possible radii of these cluster, corresponding to a more and less clumped
point distribution (radius = 5 and 15% of the extent of the landscape). Each scenario is simulated with 10, 30, 50, 100, 200, and 1000 points, so that we have 48 scenarios in total.

```{r several_point_patterns}
set.seed(1234)

# random, gradient, single center, multiple centers
methods <- c("regular", "random", "mobsim")
name <- c("regular", "random", "clumped5", "clumped1")
nfeat <- c(10, 30, 50, 100, 200, 1000) # number of features
res <- 100 # resolution
ext <- 30000 # extent of the landscape
nc <- c(5, 1) # number of centers or clusters for clumped
wd <- c(0.05, 0.15)*ext # width of the "clusters"

# parameters
parms_df1 <- expand.grid(method = methods, n_features = nfeat,
                        centers = nc[1], width = wd) # first 3 scenarios
parms_df2 <- expand.grid(method = methods[3], n_features = nfeat,
                         centers = nc[2], width = wd) # parameters for clumped1
parms_df <- dplyr::bind_rows(parms_df1, parms_df2) %>% 
  dplyr::arrange(width, n_features, method)
# names of scenarios
scenarios <- paste0(rep(name, 12), "_", rep(rep(nfeat, each = 4), 2), "_", rep(wd/300, each = 24))  

# simulate points
pts <- parms_df %>% purrr::pmap(set_points, res = res, 
                                extent_x = c(0, ext),
                                extent_y = c(0, ext), 
                                buffer_around = 10000)

landscapes <- purrr::map(pts, ~ .[[2]])
names(landscapes) <- scenarios
```

As an example, here we visualize the scenarios with 1000 features, both for clusters with small and intermediate sized radius (5% and 15% of the landscape extent).

```{r visualize_1000}
# show landscapes with n_features = 1000
# plot(landscapes, col = "black", nc = 4, legend = F)
# rasterVis::levelplot(stack(landscapes), layout = c(4,3), names.attr = scenarios,
#                      par.settings = GrTheme, colorkey = FALSE)


landscapes[c(21:22, 47:48, 23:24)] %>% 
  purrr::map(raster::crop, y = extent(0, ext, 0, ext)) %>% 
  landscapetools::show_landscape()
```

We also store some variables related the distance between points for each of these scenarios. These variables are: (i) the average distance between points, (ii) the average nearest neighbor distance, and (iii) the average isolation. The average isolation is defined here as the mean nearest neighbor distance between random points created in the landscapes and the simulated point feature locations.

```{r dist_points_patterns}
# points
pts_coords <- purrr::map(pts, first)

# calculate distances
dist_scenarios <- data.frame(
  scenario = scenarios,
  spatial_dist = rep(name, 3),
  n_features = rep(nfeat, each = 4),
  cluster_radius = rep(wd/300, each = 24),
  mean_dist = map_dbl(pts_coords, function(x) mean(dist(x))),
  mean_nndist = map_dbl(pts_coords, function(x) mean(spatstat.geom::nndist(x))),
  mean_isolation = map_dbl(pts_coords, mean_isolation, n_rand = 150, ext = c(0, ext))
)

head(dist_scenarios)
```

Just to understand how these distance and isolation measures change among scenarios, we plot them below, for both small and large clusters for the clumped scenarios (radius = 5% and 15% of the landcape extent).

```{r scenarios_dist_patterns}
# relationship between scenarios
ds1 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_dist, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  facet_wrap(~cluster_radius, ncol = 1) +
  ylim(0, NA) +
  labs(x = "",
       y = "Mean distance",
       color = "N. features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
ds2 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_nndist, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  facet_wrap(~cluster_radius, ncol = 1) +
  ylim(0, NA) +
  labs(x = "Spatial distribution",
       y = "Mean nearest neighbor dist.",
       color = "N. features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
ds3 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_isolation, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  facet_wrap(~cluster_radius, ncol = 1) +
  ylim(0, NA) +
  labs(x = "",
       y = "Mean isolation",
       color = "N. features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

# combine
ggpubr::ggarrange(ds1, ds2, ds3, nrow = 1, common.legend = T, 
                  legend = "right")
```

As expected, the mean distance between points and the mean distance to the nearest neighbor
point increase when we depart from a more clumped scenario to a random or regular distribution.
They represent the distance relationship between points only, regardless of the size of the landscape

The mean isolation follows an opposite pattern: it decreases when we depart
from clumped to random or regular distribution of points. In all cases, even though the general
pattern is qualitatively the same, quantitatively these quantities vary with the number of
features - obviously the distances or isolation are larger when the density of points is smaller.

We take the last measure - mean isolation - to represent the relationship between
the spatial distribution of points and the correlation between log-distances and cumulative influence
(density) measures.


## Calculate distance and cumulative influence

First we illustrate, for one of those scenarios, how the maps change as we calculate either the distance to the nearest feature or the CumInf (or density) of features at different scales. For illustration purposes, we use here a Gaussian filter to calculate CumInf, where the scale corresponds to the standard deviation $\sigma$ of the Gaussian distribution, and the moving window has a size of $3 \cdot \sigma$.

```{r one_dist_dens, message=FALSE}
# calculate distance and density at multiple scales for one input
scales <- c(250, 500, 1000, 2500, 5000)/2 # scales for Gaussian filter
dist_dens1 <- calc_dist_dens(landscapes[[2]], type_density = "Gauss", scale = scales,
                             extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))
plot(dist_dens1, nc = 3)
```


We can see that the results are different if one transforms the distance variable, as it is generally done in ecological models. Below we show, as an example, how it looks like when the distances are log- and sqrt-transformed, or when an exponential decay distance is used instead. 
<!-- It also affects how distances and densities correlate, as we'll see. -->

```{r log_trasnform_dist}
# transformed distance
log_dist <- calc_dist(landscapes[[2]], transform_dist = "log", log_base = 10,
                      extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))
sqrt_dist <- calc_dist(landscapes[[2]], transform_dist = "sqrt",
                       extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))
exp_decay_dist <- calc_dist(landscapes[[2]], transform_dist = "exp_decay",
                       exp_hl = 1000,
                       extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))

# combine
dists <- raster::stack(dist_dens1[[1]], log_dist, sqrt_dist, exp_decay_dist)

# plot
names(dists) <- c("dist", "log-dist", "sqrt-dist", "exp-decay-dist")
plot(dists, nc = 2)
```

## Visual comparison

Now we compare these variables visually for the different spatial point patterns, using the log-transformed distance, for now. We select the log-distance since it is
a measure commonly used in ecological studies to include distance to the nearest features into the statistical models.

```{r several_dist_dens, message=FALSE}
# redefine scales
scales <- c(250, seq(500, 5000, by = 500))/2 # scales for Gaussian filter

# calculate distance and density at multiple scales for each input
dist_dens <- purrr::map(landscapes, calc_dist_dens,
                        type_density = "Gauss", scale = scales,
                        # type_density = "circle", scale = scales,
                        transform_dist = "log", log_base = 10,
                        extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))

# change names according to the scenario and the variable
vars <- c("logdist", paste0("dens", scales*2))

for(i in 1:length(dist_dens)) 
  names(dist_dens[[i]]) <- paste(strsplit(names(dist_dens)[i], "_")[[1]][1], vars, sep = "_")

# slice a few of them to plot - for 
slices <- c(1,2,4,6) #as.vector(outer(c(1,2,4,6), (0:3)*6, FUN = "+"))
maps100 <- dist_dens[13:16]
stk <- raster::stack(maps100[[1]][[slices]], maps100[[2]][[slices]], 
                     maps100[[3]][[slices]], maps100[[4]][[slices]])
plot(stk, nc = 4)
```

## Calculate correlations between distance and cumulative influence

Now we are going to check how correlated these variables are in space, to evaluate how their ecological interpretations might differ.

```{r correlation1}
# put distances with rasters
scenarios_analysis_full <- dist_scenarios %>% 
  tibble::as_tibble() %>% 
  dplyr::mutate(rasts = dist_dens,
                # rasters as data.frame
                rasts_df = purrr::map(rasts, as.data.frame),
                # correlation between pairs of rasters
                corr_raw = purrr::map(rasts_df, lsr::correlate),
                # select only correlation between log-dist and CumInf at different ZoI
                corr_list = purrr::map(corr_raw, function(x) x$correlation[1, 2:12]),
                # transform into tibble
                corr_df = purrr::map(corr_list, tibble::enframe, name = "dens", value = "corr")) %>% 
  tidyr::unnest(corr_df) %>% 
  # organize names
  dplyr::mutate(variable = gsub(".*_", "log-dist_", dens),
                scale = as.numeric(gsub("\\D+", "", variable)))

# select only relevant variables for plotting
scenarios_analysis <- scenarios_analysis_full %>% 
  dplyr::select(-c(rasts:corr_list)) %>% 
  dplyr::mutate(n_features = factor(n_features))
  
# plot
dat_isol <- scenarios_analysis %>% 
  dplyr::filter(cluster_radius == 5, spatial_dist %in% c("random", "regular")) %>% 
  dplyr::group_by(n_features, spatial_dist) %>% 
  dplyr::summarise(mean_iso = unique(mean_isolation))

scenarios_analysis %>% 
  dplyr::filter(cluster_radius == 5) %>% 
  ggplot(aes(scale, abs(corr), color = spatial_dist)) +
  geom_point() +
  geom_line() +
  geom_vline(data = dat_isol, aes(xintercept = mean_iso, linetype = spatial_dist)) +
  ylim(0, 1) +
  facet_wrap(~n_features) +
  labs(x = "Zone of Influence", 
       y = "Absolute correlation",
       color = "Spatial distribution",
       linetype = "Mean isolation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

We plot here the the absolute value of the Pearson correlation between values of the log-distance
and the CumInf at mutltiple scales/ZoI, and highlight in each plot the average isolation between
points in the random and regular scenarios (vertical lines). We use the absolute correlation because we
are interested in the degree of correlation but not in the signal of the correlation.
The plot above is for small clusters of point features (radius = 5% of the study area extent).

We can see that the pattern changes greatly with both the total number of features 
and their spatial distribution. 

For clumped distribution of features and when the cluster of features are small 
(as in the figure above), the point features are in general quite apart from each other,
so as the ZoI increases the cumulative influence (density) measure gets 
more correlated to the log-distance, regardless of the density of features in space.

That is also true for regular and random distribution of features when the density 
of features is small (n = 10) and, consequently, their mean isolation is also high.
As the density of features increase and the mean isolation between points 
decreases, the CumInf (density) measure starts to present a peak when 
ZoI is close to the mean isolation between features in the landscape - compare
the peaks in the plots above with the mean isolation for the random and regular
scenarios. For larger ZoI values, the correlation drops very 
quickly, especially for the regular distribution scenario.

This means that, when comparing log-distances to CumInf 
for random and regular distribution of features,
they only represent similar spatial variation when the the mean isolation of the
features is around the same size of their ZoI. This is also true for clumped 
distribution when the radius of the clusters of features is larger, such as
in the figure below (radius = 15% of the landscape extent). 

When the features are distributed in more than one cluster, 
the correlation between log-distance and the
cumulative influence measure also starts to show a peak related to the average
distance between features. As this radius increases, the average isolation between 
features decreases and the results of the clumped distributions
gets closer to the one for random distribution.

The only case when this changes is when there is a single small clump of features.

```{r correlation2}
# plot
dat_isol <- scenarios_analysis %>% 
  dplyr::filter(cluster_radius == 15, spatial_dist %in% c("clumped5")) %>% 
  dplyr::group_by(n_features, spatial_dist) %>% 
  dplyr::summarise(mean_iso = unique(mean_isolation))

scenarios_analysis %>% 
  dplyr::filter(cluster_radius == 15, spatial_dist %in% c("clumped5", "clumped1")) %>% 
  ggplot(aes(scale, abs(corr), color = spatial_dist)) +
  geom_point() +
  geom_line() +
  geom_vline(data = dat_isol, aes(xintercept = mean_iso, linetype = spatial_dist)) +
  ylim(0, 1) +
  facet_wrap(~n_features) +
  labs(x = "Zone of Influence", 
       y = "Absolute correlation",
       color = "Spatial distribution",
       linetype = "Mean isolation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
<!-- Now we use, instead of the name of the scenario, the average distance measures  -->
<!-- taken for each, to see if any response can be better interpreted in this terms.  -->
<!-- First, below we see how these variables change for each scenario and density of  -->
<!-- features. -->

```{r correlation3, eval=FALSE, include=FALSE}
# relationship between scenarios
ds1 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_dist, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  ylim(0, NA) +
  labs(x = "",
       y = "Mean distance",
       color = "N. features") +
  theme(axis.text.x = element_blank())
ds2 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_nndist, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  ylim(0, NA) +
  labs(x = "Spatial distribution",
       y = "Mean nearest neighbor dist.",
       color = "N. features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
ds3 <- dist_scenarios %>% 
  ggplot(aes(x = spatial_dist, y = mean_isolation, color = factor(n_features))) +
  geom_point() +
  geom_line(aes(group = n_features)) +
  ylim(0, NA) +
  labs(x = "Spatial distribution",
       y = "Mean isolation",
       color = "N. features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

# combine
ggpubr::ggarrange(ds1, ds2, ds3, nrow = 2, ncol = 2, common.legend = T, 
                  legend = "right")
```

<!-- As expected, the mean distance between points and the mean distance to the nearest neighbor -->
<!-- point increase when we depart from a more clumped scenario to a random or regular distribution. -->
<!-- They represent the distance relationship between points only, regardless of the size of the landscape -->

<!-- The mean isolation follows an opposite pattern: it decreases when we depart  -->
<!-- from clumped to random or regular distribution of points. In all cases, even though the general -->
<!-- pattern is qualitatively the same, quantitatively these quantities vary with the number of -->
<!-- features - obviously the distances or isolation are larger when the density of points is smaller. -->

<!-- We take the last measure - mean isolation - to represent the relationship between  -->
<!-- the spatial distribution of points and the correlation between log-distances and cumulative impact -->
<!-- (density) measures. -->

```{r correlation4, eval=FALSE, include=FALSE}
scenarios_analysis %>% 
  # dplyr::filter(n_features == 1000) %>% 
  ggplot(aes(scale, abs(corr), color = mean_isolation)) +
  geom_point() +
  ylim(0, 1) +
  labs(x = "Zone of Influence", 
       y = "Absolute correlation",
       color = "Mean isolation") +
  facet_wrap(~n_features) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```


## Discussion

Here we have compared how different is the spatial variation when one computes
log-distances and cumulative influences (or densities) from point infrastructure. 
To do so, we simulated a different number of point
features in space spread according to different and spatial distributions, calculated 
the log-distances and CumInf at multiple scales (ZoI) and compared them
through correlation.

We found the correlation between log-distance and CumInf is intrinsically linked
with the mean isolation of the point features in space.

When all point infrastructure features are limited to a single, small cluster, 
(e.g. one single village or wind park in the landscape)
so that their mean isolation is very high,
log-distance and CumInf are little correlated, and their correlation increases as 
the scale (ZoI) increases. In this case, they start to represent a more similar
variation when the scale/ZoI is high. This also holds true when there more then one
cluster of points but their radius is relatively small 
(scenario with 5 small clumps shown above),
a case in which the mean isolation is still high.

In all other cases - either a regular or random distribution of points, or a clumped
distribution with more clusters or larger clusters - the log-distance is only highly
correlated with the CumInf (density) when the scale (ZoI) is at the same order of magnitude
of the mean isolation to the point features.

Given that the determination of the scale/ZoI is an empirical question and varies
according to the ecological system and context of the study area, we argue here 
that in very limited cases 
using the distance to the nearest feature might be equivalent to using CumInf or 
the density of infrastructure. In all other cases, the CumInf might represent a 
wider range of spatial variation metrics, that in turn should better represent
the cumulative effect of the multiple infrastructure in space.

**ADD comment here about the distance measure. For sqrt dist, the results are qualitatively similar. For exp decay, please see the other sup material**