---
title: "Appendix A"
subtitle: "Defining and deriving the zone of influence for multiple features"
author: ""
date: ""
geometry: margin=2cm
abstract: | 
  In this document, we provide additional details around the definition of the zone of influence (ZOI). First, we define the ZOI $\phi$ and present examples of ZOI functions with different shapes, parameterized with the ZOI radius. Second, we derive the ZOI of the nearest 
  feature and the cumulative ZOI of multiple features of an infrastructure for types of infrastructure and spatial variables that are
  represented as lines (e.g. roads, power lines) or polygons (e.g. dams, mining sites). This complements the derivation for point-type
  features as presented in the main text of Niebuhr et al. *Estimating the cumulative impact and zone of influence of anthropogenic
  features on biodiversity.* Methods in Ecology and Evolution.
bibliography: cuminf_bib.bib
output:
  NinaR::jensAnalysis:
  highlight: tango
  fig_caption: yes
  toc: yes
header-includes:
  \usepackage{float}
  \usepackage{caption}
  \DeclareCaptionLabelFormat{Sformat}{#1 A#2}
  \captionsetup[figure]{labelformat=Sformat}
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
library(NinaR)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/",
  fig.align="center",
  fig.pos = 'H',
  fig.width = 5,
  fig.asp = 0.7,
  message=FALSE,
  warning=FALSE,
  options(scipen=999)
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())
```

# Introduction

In the main text, we defined the zone of influence (ZOI) and derived the estimation 
of the ZOI of multiple features of an infrastructure type under different 
assumptions, but only did so for anthropogenic features that can be represented as points, 
such as houses, tourist cabins, or wind turbines. Here we complement it by exemplifying 
ZOI functions with different shapes, parameterized on the ZOI radius $r$, and derive 
similar equations for linear infrastructure (e.g. roads, railways, trails, power lines) 
and infrastructure or landscape variables represented by polygons (e.g. mining sites, 
hydro dams, deforestation areas or polygons outlining specific land cover and land use types).

# Definition of the Zone of Influence

As defined in the main text, the ZOI is the function $\phi$ that informs how the 
impact of a given feature decreases with distance. Formally, the 
ZOI $\phi = f(d, r)$ is any decay function that has a maximum value 
1 where the feature is located, decreases towards zero as the Euclidean 
distance $d$ increases, and possibly vanishes at a given point, the ZOI radius $r$. 
Broadly speaking, the ZOI is characterized by its shape and radius. Here we use 
four functions with different shapes to evaluate how the impact decays with distance, 
even though in principle any function $\phi$ as defined above could be used. 
Below we write these four functions and show how the radius $r$ is defined in each one.

## Functions with a well-defined ZOI radius

Some functions vanish for a certain non-infinite distance and therefore present 
well-defined radii. Here the ZOI radius $r$ represents the distance beyond which $\phi = 0$.
We used two functions within this class: the threshold and the linear decay (Bartlett)
functions.

### Threshold function

The threshold function is defined as 

$$
\phi_{threshold}(d_{i,k}, r_k) = 
    \begin{cases}
      1 & \text{if $d_{i,k} < r_k$} \\
      0 & \text{if $d_{i,k} \ge r_k$}
    \end{cases}
$$

where $d_{i,k}$ is the distance to the feature $i$ of type $k$ and $r_k$ 
is the radius of the ZOI for the features of type $k$. Then, $\phi_{threshold}$ 
represents a constant impact of each feature until $d = r$ and is 
similar to what has generally been used when buffer zones are created around 
anthropogenic features to calculate derived measures.

Fig. A\ref{fig:plot_threshold} shows a plot of $\phi_{threshold}$ with $r = 10$ km.

```{r plot_threshold, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.cap="Illustration of a threshold ZOI function with radius = 10 km, for a feature located at the origin."}
library(oneimpact)
library(ggplot2)

oneimpact::plot_zoi1d(points = 0, radius = 10, fun = threshold_decay, 
                      range_plot = c(-20, 20)) +
  labs(x = "Distance (km)")
```

### Linear decay function

The linear decay function, also called tent decay function or "Bartlett" filter 
(Harris, 1978), corresponds to 1 at the feature location and decreases 
linearly with the distance from it, until it reaches zero at a given distance,
defined here as the ZOI radius $r$. Formally, the linear ZOI function 
$\phi_{linear}$ is defined as:

$$
\phi_{linear}(d_{i,k}, r_k) = 
    \begin{cases}
      1 - \frac{d_{i,k}}{r_k} & \text{if $d_{i,k} < r_k$} \\
      0 & \text{if $d_{i,k} \ge r_k$}
    \end{cases}
$$

where $d_{i,k}$ is the distance to the feature $i$ of type $k$ and $r_k$ 
is the radius of the ZOI for features of type $k$.

Fig. A\ref{fig:plot_linear} shows a plot of $\phi_{linear}$ with $r = 10$ km.

```{r plot_linear, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.cap="Illustration of a linear decaying ZOI function with radius = 10 km, for a feature located at the origin."}
oneimpact::plot_zoi1d(points = 0, radius = 10, fun = linear_decay, 
                      range_plot = c(-20, 20)) +
  labs(x = "Distance (km)")
```

## Functions that do not vanish with distance

Some functions decrease but do not vanish as the distance from anthropogenic features increase. 
In these cases we define the ZOI radius $r$ as the distance at which the ZOI decreases
to $\phi = \phi_{limit}$, an arbitrary small ZOI value beyond which the influence
of the feature is considered to be negligible. In these cases, the ZOI 
definition needs an extra parameter and is defined as $\phi = f(d, r, \phi_{limit})$.
We used two functions within this class: the exponential decay and the Gaussian decay
functions. For both we set $\phi_{limit} = 0.05$ for all features.

### Exponential decay function

The exponential decay function is defined as 

$$
\phi_{exp}(d_{i,k}, r_k, \phi_{limit_k}) = \exp(-\lambda_{r_k} d_{i,k})
$$

where $d_{i,k}$ is the distance to the feature $i$ of type $k$ and the 
decay parameter $\lambda_{r_k}$ is defined in terms of $r_k$ and $\phi_{limit_k}$:

$$
\lambda_{r_k} = \frac{\ln(1/\phi_{limit_k})}{r_k}
$$

Here $\ln$ is the logarithm with base $e$. Fig. A\ref{fig:plot_exponential} shows 
a plot of $\phi_{exp}$ with $r = 10$ km.

```{r plot_exponential, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.cap="Illustration of an exponential decaying ZOI function with radius = 10 km, for a feature located at the origin."}
oneimpact::plot_zoi1d(points = 0, radius = 10, fun = exp_decay, 
                      range_plot = c(-20, 20)) +
  labs(x = "Distance (km)")
```

### Gaussian decay function

The Gaussian (or half-normal) decay function is defined as 

$$
\phi_{Gauss}(d_{i,k}, r_k, \phi_{limit_k}) = \exp(-\lambda_{r_k} d_{i,k}^2)
$$

where $d_{i,k}$ is the distance to the feature $i$ of type $k$ and the 
decay parameter $\lambda_{r_k}$ is defined in terms of $r_k$ and $\phi_{limit_k}$:

$$
\lambda_{r_k} = \frac{\ln(1/\phi_{limit_k})}{r_k^2}
$$

In Fig. A\ref{fig:plot_gaussian} we visualize $\phi_{Gauss}$ with $r = 10$ km.

```{r plot_gaussian, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, fig.cap="Illustration of a Gaussian decaying ZOI function with radius = 10 km, for a feature located at the origin."}
oneimpact::plot_zoi1d(points = 0, radius = 10, fun = gaussian_decay, 
                      range_plot = c(-20, 20)) +
  labs(x = "Distance (km)")
```

Notice that, even though the ZOI radius $r$ is defined for all the
functions, the change in their shape modifies the interpretation of how
the ZOI changes with distance (see also Fig. 1 in the main text).

# Derivation of zones of influence for multiple anthropogenic features of a given type

We start from the same baseline presented in the main text: we use as an example 
of statistical model a habitat-selection function (HSF; Fieberg et al., 2021). 
Copying here eq. 2 from the main text, the HSF is defined as:

\begin{equation} 
\centering
\label{eqn:HSF}
w(\textbf{X}) = \exp \left( \beta_0 + \overbrace{\beta_1 X_1}^\text{A) Infrastructure type 1} + \overbrace{\beta_2 X_2}^\text{B) Infrastructure type 2} + \underbrace{\beta_{12} X_1 X_2}_\text{D) Interaction infrastructure types 1 and 2} + ... + \overbrace{\beta_k X_k}^\text{C) Infrastructure type k} \right)
\end{equation}

where $\textbf{X} = X_1,X_2, ...,  X_k$ correspond to $k$ different types of 
infrastructure or other anthropogenic features and $\beta_k$ represents the 
effect size of the impact of spatial features of type $k$.

## ZOI metrics

As shown in the main text, when multiple features of the same type are
present in the landscape, we can calculate two ZOI metrics: the ZOI of the nearest 
feature only ($\phi_{nearest}$) and the cumulative ZOI of multiple features
($\phi_{cumulative}$).

We assume that in the landscape there are $n_k$ features of the same type $k$, 
and let the influence of the feature $i$ of a feature 
\textit{k} be a ZOI function $\phi$ as defined above. As we show in the main text, 
for point-like infrastructure we can sum the effect of each feature, so that the 
impact terms in equation \ref{eqn:HSF} become:

\begin{equation}
\label{eqn:HSFtermPoint}
    I_k = \beta_k X_k = \sum_{i=1}^{n_k} \beta_{i,k} \phi_{i_k}.
\end{equation}

Likewise, for linear infrastructure we can integrate the term

\begin{equation}
\label{eqn:HSFtermLine}
    I_k = \beta_k X_k = \int_{L} \beta_{k}(l) \ \phi_{k}(l; d_{i,k}, r_k) dl
\end{equation}

where $dl$ denotes the length on an infinitesimal infrastructure segment and $L$ is the
total length of linear infrastructure of type $k$ in the study area. 
For polygon variables we can similarly integrate the term, but on two dimensions $(x, y)$:

\begin{equation}
\label{eqn:HSFtermPolygon}
    I_k = \beta_k X_k = \iint_{\Omega} \beta_{k}(x,y) \ \phi_{k}(x, y; d_{i,k}, r_k) \ dx \ dy
\end{equation}

where $\Omega$ is the whole study area.

## Zone of influence of the nearest feature

To estimate the ZOI of the nearest feature alone, we showed that, for point-like 
features, one generally considers $\beta_i = 0$ for all $i > 1$ (where the 
features are ordered by increasing distance), i.e., only the nearest feature is 
assumed to influence a given site. We can make equivalent assumptions for linear 
and area infrastructure by considering only the nearest (orthogonal) segment of 
linear infrastructure or the closest edge of the closest polygon to influence 
processes in a given location. In such a way, their influence is similar to that 
of a point. As a consequence, eq. 4 from the main text keeps valid to represent 
the influence of the nearest feature for linear and polygon variables:

\begin{equation}
\label{eqn:HSFnearest}
\begin{split}
    I_k & = \beta_{1,k} \cdot \phi_{1,k} \\
        & = \beta_{k} \cdot \phi_{nearest_k}.
\end{split}                
\end{equation}

Here $\phi_{nearest}$ can be set to represent the distance to the nearest feature
(or some transformation of that, such as the log-distance to the nearest feature),
as it is most common in the literature. However, our formulation with the 
definition of the set of functions $\phi$ to represent the ZOI allow for more a 
general representation of the impact of the nearest feature, while still allowing 
the estimation of the shape and radius of the zone of influence.

## Cumulative zone of influence

To derive the cumulative ZOI of multiple features of a given type, 
for point-like features we assumed that all features exerted the same influence, 
so that $\beta_{1,k} =\beta_{2,k} = ... = \beta_k = \beta$. If we do similar assumptions 
for lines and polygons -- that all infinitesimal segments $dl$ present the same 
influence over $L$ and that all pieces of area of the polygons present similar 
influence around them, eq. 5 from the main text remains valid for these types of variables:

\begin{equation}
\label{eqn:HSFcumzoi}
\begin{split}
    I_k & = \beta_k \int_{L} \phi_{k}(l) \ dl \\
        & = \beta_k \ \phi_{cumulative_k} \\
        & \text{and} \\
    I_k & = \beta_k \iint_{\Omega} \phi_{k}(x, y) \ dx \ dy \\
        & = \beta_k \ \phi_{cumulative_k}.
\end{split}
\end{equation}

The only difference here regards how the cumulative ZOI function $\phi_{cumulative}$ 
is defined -- as a summation for point features (eq. 5 in the main text) or 
as an integral over 1 or 2 dimensions for linear and area features (eq. 
\ref{eqn:HSFcumzoi}). 

# Computation of the cumulative zone of influence

In theory, the calculation of the cumulative zone of influence function 
$\phi_{cumulative}$ for a given study area $\Omega$ requires the computation of 
eq. 4 from the main text or eq. \ref{eqn:HSFcumzoi} for each infrastructure feature, 
which might be difficult and computationally challenging for large areas with many 
anthropogenic features spread non-regularly in space. To overcome that limitation, 
we implemented this calculation by constructing filters (or weighing matrices; 
see Miguet et al., 2017) that follow an assumed decaying function (such as the 
ZOI functions described above) and used them on neighborhood analyses over the 
input infrastructure layers (see the [vignettes](https://ninanor.github.io/oneimpact/articles/) 
of the `oneimpact` R package for 
an illustration of how this approach is applied). This is analogous to the 
approaches presented by Miguet et al. (2017) and Gilleland (2013). Even though 
we do not demonstrate it here, for non-vanishing function (e.g. exponential or 
Gaussian decay) when the size of the filter is set properly 
to include most the area within which $\phi > 0$ (i.e. $\phi_{limit}$ is 
small enough), the result of applying eq. 4 of the main text or eq. 6 presents 
a correlation $> 0.99$ with the result of the neighborhood analysis.

# References

Fieberg, J., Signer, J., Smith, B., & Avgar, T. (2021). A 'How to' guide for 
interpreting parameters in habitat‐selection analyses. Journal of Animal Ecology, 
90(5), 1027--1043. https://doi.org/10.1111/1365-2656.13441

Gilleland, E. (2013). Two-dimensional kernel smoothing: Using the R package smoothie. 
NCAR Technical Note, TN-502+STR, 17pp. https://doi.org/10.5065/D61834G2

Harris, F. J. (1978). On the use of windows for harmonic analysis with the discrete 
Fourier transform. Proceedings of the IEEE, 66(1), 51–83. 
https://doi.org/10.1109/PROC.1978.10837

Miguet, P., Fahrig, L., & Lavigne, C. (2017). How to quantify a distance‐dependent 
landscape effect on a biological response. Methods in Ecology and Evolution, 8(12), 
1717--1724. https://doi.org/10.1111/2041-210X.12830

