---
title: "Appendix 4"
subtitle: "Cumulative impacts of infrastructure on reindeer space use: calculating the ZoI metrics"
author: "Bernardo Niebuhr, Bram van Moorter, Manuela Panzacchi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=2cm
abstract: | 
  In this document, we prepare covariates and wild reindeer movement data for habitat selection 
  analysis presented in Niebuhr et al. *Estimating the cumulative impact and zone of influence 
  of anthropogenic infrastructure on biodiversity.* The main goal of this document is to show
  how to calculate the cumulative ZoI and the ZoI from the nearest feature for several
  infrastructure variables using the R-GRASS interface and the `oneimpact` package.
output:
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
library(NinaR)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/",
  fig.align="center",
  message=FALSE,
  warning=FALSE
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())
```

\newpage

In this document we prepare covariates and wild reindeer movement data for the 
Hardangervidda area for habitat selection analysis. The reindeer GPS dataset is 
organized in a setup of use-availability, where used locations (column 
`use = 1`) are compared to locations available but not used by the animals 
(column `use = 0`). We annotate the GPS dataset with the ZoI metrics for the 
following covariates:

- Private cottages
- Large tourist cabins
- Small tourist cabins
- Roads of high traffic
- Roads of low traffic
- Trails

We also annotate the GPS data with:

- Land cover
- 4 PCAs corresponding to bio-climatic-geographical variation in Norway

For the infrastructure covariates, we calculate the cumulative ZoI and the ZoI 
of the nearest feature considering different ZoI functions (threshold, linear decay, 
Gaussian decay, and exponential decay) and different radii for the zone of influence, 
from 100 m to 20 km.

Below we present the steps for this data preparation.

\newpage

# Load packages

We start loading the packages necessary for the analysis. Spatial analyses are done 
through GRASS GIS using the `oneimpact` package.

```{r load_packages, eval=FALSE, message=FALSE, warning=FALSE}
#Load packages
library(NinaR)

# data manipulation
library(dplyr)
library(purrr)
library(ggplot2) # for plot
library(ggpubr) # for plot

# spatial packages
library(rgrass7)
library(terra)
library(sf)
library(tmap)

# oneimpact package
library(oneimpact)
```

# Connect to GRASS

We start by connecting the R session to GRASS GIS. We link it to the GRASS GIS 
database from the Norwegian Institute for Nature Research (NINA), where the spatial 
data is already organized.

```{r connect_grass, echo=TRUE, eval=FALSE}
# Connect to GRASS GIS location and mapset
ms <- "u_bb_cuminf"
NinaR::grassConnect(mapset = ms)

# set flags
flags <- c("quiet", "overwrite")
```

\newpage

# Set the study area

Here we subset the vector with the limits of the study area, the Hardangervidda 
wild reindeer area, and make a 50 km buffer around it, for preparing the spatial 
data considering also the surrounding of the study area and to avoid edge effects 
in the spatial data preparation.

```{r load_vectors, echo=TRUE, eval=FALSE}
# load study area
areas_g <- "wild_reindeer_areas@u_bram.van.moorter"
study_area_n <- "Hardangervidda"

# load study area
study_area <- rgrass7::read_VECT(areas_g) %>%
  sf::st_as_sf() %>% 
  dplyr::filter(name_area == study_area_n) %>%
  sf::st_union()
# plot(study_area)

# buffer around the study area
buff <- 50e3 # 50km buffer
study_area_buff50 <- study_area %>%
  sf::st_buffer(dist = buff)
# plot(study_area_buff50)

# put study area and buffer into the GRASS GIS mapset
study_area_g <- "study_area"
rgrass7::write_VECT(study_area, study_area_g,
                    v.in.ogr_flags = flags)
study_area_buff_g <- "study_area_buff50"
rgrass7::write_VECT(study_area_buff50, study_area_buff_g,
                    v.in.ogr_flags = flags)
```

Here is just a plot of the study area, for illustration.

```{r viz1, echo=TRUE, eval=FALSE}
tm_shape(study_area_buff50) +
  tm_borders() +
  tm_shape(study_area) +
  tm_borders() +
  tm_scale_bar(breaks = c(15, 30, 45, 60))
```

\newpage

# Rasterize infrastructure vectors

Now we rasterized the vector layers corresponding to the infrastructure variables we
considered for this area. We used a previously created raster file with 100 m resolution as
a base grid for rasterizing the infrastructure layers.

For the private cottage and tourist cabins, we first created rasters with value 1 where there
are infrastructure and NA where they are absent, to be used as input for the calculation
of the ZoI of the nearest feature. Here it is not necessary to count the number of features,
only to know if they are present or not in a given pixel, since this measure takes into
account of the ZoI of the nearest feature.

We also created rasters counting the number of cottages and cabins in each pixel, 
which were used as input for the calculation of the cumulative ZoI. This procedure was
made with the `util_v2rast_count_grass` function of the `oneimpact` package.

```{r rasterize_maps_count, echo=TRUE, eval=FALSE}
# set region as the study area buffer
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# vectors to be rasterized
vectors_to_rasterize_c <- c("private_cabins@p_prodchange_envpoints",
                            "pub_cabins_summer_high@p_prodchange_envpoints",
                            "pub_cabins_summer_low@p_prodchange_envpoints")
# output raster map names
outputs_rasterize_c1 <- c("private_cabins",
                          "public_cabins_high",
                          "public_cabins_low")

# option
use_rasterize_c1 <- "val"

#---
# First rasterize using val=1 to use as input for calc_zoi_nearest

# put all in a tibble
rasterize_parms_c1 <- tibble::tibble(input = vectors_to_rasterize_c,
                                     output = outputs_rasterize_c1,
                                     use = use_rasterize_c1)

# rasterize
rasterized_c1 <- rasterize_parms_c1 %>% purrr::pmap(rgrass7::execGRASS, cmd = "v.to.rast",
                                                    flags = flags)

# colors
f <- tempfile()
cat("1 grey", file = f)
# readr::read_file(f)
colored <- purrr::map(outputs_rasterize_c1, function(x)
  rgrass7::execGRASS(cmd = "r.colors", map = x, rules = f, flags = ))

#---
# Then rasterize using count to use as input for calc_zoi_cumulative

# output raster map names
outputs_rasterize_c2 <- c("private_cabins_count",
                          "public_cabins_high_count",
                          "public_cabins_low_count")

# parameters
count_parms <- tibble::tibble(x = vectors_to_rasterize_c,
                              output = outputs_rasterize_c2,
                              set_region = FALSE)

# rasterize with count
count <- count_parms %>% purrr::pmap(oneimpact::util_v2rast_count_grass,
                                     quiet = F, overwrite = TRUE)

```

For the roads and trails, however, for 100m pixels there can hardly be more than one 
feature in each pixel, and if this occurs it is not relevant for the analysis. Therefore, 
a rasterization counting the number of features was not made. Instead, besides the common
rasterization used to create the inputs for the calculation of the ZoI of
the nearest feature, we create binary raster maps (1 where there are infrastructure, 0 
elsewhere) to be used as input for the cumulative ZoI calculation.

```{r rasterize_maps_value1, echo=TRUE, eval=FALSE}
# set region as the study area buffer
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# vectors to be rasterized
vectors_to_rasterize_1 <- c("roads_summer_high@p_prodchange_roads",
                            "roads_summer_low@p_prodchange_roads",
                            "trails@p_RenRein_trails2")
# output raster map names
outputs_rasterize_1 <- c("roads_high",
                         "roads_low",
                         "trails")

# option - values = 1
use_rasterize_1 <- c("val")

# put all in a tibble
rasterize_parms_1 <- tibble::tibble(input = vectors_to_rasterize_1,
                                    output = outputs_rasterize_1,
                                    use = use_rasterize_1)

# rasterize
rasterized_1 <- rasterize_parms_1 %>% purrr::pmap(rgrass7::execGRASS, cmd = "v.to.rast",
                                                  flags = flags)

# colors
f <- tempfile()
cat("1 grey", file = f)
# readr::read_file(f)
colored <- purrr::map(outputs_rasterize_1, function(x)
  rgrass7::execGRASS(cmd = "r.colors", map = x, rules = f, flags = "quiet"))

# binarize maps for cumulative influence calculation
output_bin_1 <- paste0(outputs_rasterize_1, "_bin")

binarize_parms_1 <- tibble::tibble(x = outputs_rasterize_1,
                                   output = output_bin_1,
                                   null = 0)

binarized_1 <- binarize_parms_1 %>% purrr::pmap(oneimpact::util_binarize_grass,
                                                quiet = TRUE, overwrite = TRUE)
binarized_1 <- unlist(binarized_1)
```

```{r cut_and_save_layers, echo=FALSE, eval=FALSE}
vect_export <- c(vectors_to_rasterize_c, vectors_to_rasterize_1)
vect_export_within <- strsplit(vect_export, "@") %>% 
  sapply(first) %>% 
  paste(., "within", sep = "_")
study_area_g <- "study_area"
out_names <- gsub("within", "v.gpkg", vect_export_within)

# cut and retrieve maps for the study area
layers_in_r <- list()
for(i in 2:length(vect_export)) {
  cut
  rgrass7::execGRASS("v.select", ainput = vect_export[i], binput = study_area_g,
                     output = vect_export_within[i], operator = "within", flags = "overwrite")
  retrieve to R
  layers_in_r[[i]] <- rgrass7::read_VECT(vect_export_within[i])
  # export from GRASS
  # rgrass7::execGRASS("v.out.ogr", input = vect_export_within[i], 
  #                    output = paste0("../data/analysis_GPS/", out_names[i]))
}

# export from R
for(i in 2:length(out_names)) {
  terra::writeVector(layers_in_r[[i]], filename = paste0("../data/analysis_GPS/", out_names[i]), 
                     overwrite = TRUE)
}
```

\newpage

# Calculate the zone of influence of the nearest feature

Here we calculated, for each infrastructure type, the ZoI of the nearest feature,
for different ZoI decay functions and ZoI radii. To do so, we used the function
`calc_zoi_nearest` from the `oneimpact` package. We also calculated the log-distance
from the nearest feature, which is commonly used in ecological studies, to compare its
performance in model fitting with the other ZoI variables.

```{r calculate_nearest, echo=TRUE, eval=FALSE}
# calculate ZoI from the nearest feature
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# define parameters
nearzoi_parms <- expand.grid(zoi = zoi_vals, type = shapes)

# calculate ZoI from the nearest feature

# private cabins
nearzoi_private_cabins <- nearzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                        x = "private_cabins",
                                                        where = "GRASS",
                                                        quiet = TRUE, overwrite = TRUE)

# public cabins high
nearzoi_public_cabins_high <- nearzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                            x = "public_cabins_high",
                                                            where = "GRASS",
                                                            quiet = TRUE, overwrite = TRUE)

# public cabins low
nearzoi_public_cabins_low <- nearzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                           x = "public_cabins_low",
                                                           where = "GRASS",
                                                           quiet = TRUE, overwrite = TRUE)

# roads high
nearzoi_roads_high <- nearzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                    x = "roads_high",
                                                    where = "GRASS",
                                                    quiet = TRUE, overwrite = TRUE)

# roads low
nearzoi_roads_low <- nearzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                   x = "roads_low",
                                                   where = "GRASS",
                                                   quiet = TRUE, overwrite = TRUE)

# summer trails
nearzoi_trails <- nearinf_parms %>% purrr::pmap(oneimpact::calc_zoi_nearest,
                                                x = "trails",
                                                where = "GRASS",
                                                quiet = TRUE, overwrite = TRUE)

# additionally, also calculate log-distance
infrastructure <- c("private_cabins", "public_cabins_high", "public_cabins_low", 
                    "roads_high", "roads_low", "trails")

# calculate log-distance from the nearest feature for all infrastructure
nearzoi_log <- purrr::map(infrastructure, oneimpact::calc_zoi_nearest,
                          type = "log", where = "GRASS",
                          quiet = TRUE, overwrite = TRUE)
```

```{r calc_save_layers_euc_dist, echo=FALSE, eval=FALSE, fig.show='hide', results='hide'}
# infra
infrastructure <- c("private_cabins", "public_cabins_high", "public_cabins_low", 
                    "roads_high", "roads_low", "trails")

# region
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# calculate Euclidean distance
nearzoi_euc <- purrr::map(infrastructure, oneimpact::calc_zoi_nearest,
                          type = "euclidean", where = "GRASS",
                          quiet = TRUE, overwrite = TRUE)
nearzoi_euc <- unlist(nearzoi_euc)

# cut and retrieve maps for the study area
layers_in_r <- list()
for(i in 1:length(nearzoi_euc)) {
  # retrieve to R
  layers_in_r[[i]] <- rgrass7::read_RAST(nearzoi_euc[i], return_format = "terra")
  #export from R
  terra::writeRaster(layers_in_r[[i]], 
                     filename = paste0("../data/analysis_GPS/", nearzoi_euc[i], ".tif"), 
                     overwrite = TRUE)
  # export from GRASS
  # rgrass7::execGRASS("v.out.ogr", input = vect_export_within[i], 
  #                    output = paste0("../data/analysis_GPS/", out_names[i]))
}

plot(do.call("c", layers_in_r))
```

\newpage

# Calculate the cumulative ZoI of multiple features

Here we calculated, for each infrastructure type, the cumulative ZoI of multiple features,
for different ZoI functions and radii. To do so, we used the function
`calc_zoi_cumulative` from the `oneimpact` package.

```{r calculate_cumulative, echo=TRUE, eval=FALSE}
# calculate cumulative ZoI

# input maps, ZoI values, and filter shapes
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# define parameters
cumzoi_parms <- expand.grid(zoi = zoi_vals, type = shapes)

# calculate cumulative zoiluence

# private cabins
cumzoi_private_cabins <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                                      x = "private_cabins_count",
                                                      where = "GRASS",
                                                      quiet = TRUE, overwrite = TRUE)

# public cabins high
cumzoi_public_cabins_high <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                                          x = "public_cabins_high_count",
                                                          where = "GRASS",
                                                          quiet = TRUE, overwrite = TRUE)

# public cabins low
cumzoi_public_cabins_low <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                                         x = "public_cabins_low_count",
                                                         where = "GRASS",
                                                         quiet = TRUE, overwrite = TRUE)

# roads high
cumzoi_roads_high <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                                  x = "roads_high_bin",
                                                  where = "GRASS",
                                                  quiet = TRUE, overwrite = TRUE)

# roads low
cumzoi_roads_low <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                                 x = "roads_low_bin",
                                                 where = "GRASS",
                                                 quiet = TRUE, overwrite = TRUE)

# summer trails
cumzoi_trails <- cumzoi_parms %>% purrr::pmap(oneimpact::calc_zoi_cumulative,
                                              x = "trails_bin",
                                              where = "GRASS",
                                              quiet = TRUE, overwrite = TRUE)

```

\newpage

# Prepare and annotate GPS data

Finally, we loaded the reindeer GPS data, prepared it for analysis, and annotated 
it with the infrastructure data to allow assessing their effects on reindeer space use.

The reindeer data was already organized in a use-availability design, where used locations (column 
`use = 1`) are compared to locations available but not used by the animals (column `use = 0`). It 
was also already annotated with land cover classes (that were recombined here) and the PCA climatic
variables.

```{r read_gps, echo=TRUE, eval=FALSE}
# load gps data
filename <- "wrein_dat_sub_20200518V2.rda"
load(file=paste0("/data/R/Prosjekter/Rein/RSF_output/", filename)) #servers, e.g. ninrstudio03
study_area_n <- "Hardangervidda"

# subset summer
dat <- dat$sum
names(dat)
dat <- dat[dat$herd == study_area_n,] 

# reclass land cover
dat$NORUTreclass <- ifelse(dat$norut<9, "11forest", ifelse(dat$norut<12, "bog", 
                                                           ifelse(dat$norut<21, "mountain", 
                                                                  ifelse(dat$norut<22, "glacier", 
                                                                         ifelse(dat$norut<23, "water", 
                                                                                ifelse(dat$norut<25,"other", NA))))))
dat$NORUTreclass <- ifelse(dat$NORUTreclass=="mountain", dat$norut, dat$NORUTreclass)
tapply(dat$use, dat$NORUTreclass, mean) # explore the proportion of each used point in classes
tapply(dat$use, dat$NORUTreclass, length) # explore the number of points in each class
sum(is.na(dat$NORUTreclass))

# remove old infrastructure columns
to_remove <- c("houses", "powerlines", "private_cab", "pub_cab", 
               "railway", "roads", "ski", "trails") %>%
  paste(collapse = "|")
dat <- dat %>%
  dplyr::select(-c(grep(to_remove, colnames(dat))))
names(dat)
```

Finally, we annotated the GPS dataset with the influence measures calculated in this 
document and saved it externally for the statistical analysis.

```{r annotate_gps, echo=TRUE, eval=FALSE}
# variables
infrastructure = c("private_cabins", "public_cabins_high", "public_cabins_low", 
                   "roads_high", "roads_low", "trails")
measures <- c("cumulative", "nearest")
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000)#, 30000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# parameters for the layers
layers_parms <- expand.grid(infrastructure = infrastructure,
                            measure = measures,
                            shape = shapes,
                            zoi = zoi_vals) %>%
  dplyr::arrange(infrastructure, measure, shape, zoi)
# add log distance to parameters
layers_parms <- layers_parms %>% 
  dplyr::bind_rows(
    expand.grid(infrastructure = infrastructure,
                measure = "nearest",
                shape = "log",
                zoi = NA))

# re-set region
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# list all relevant layers
layers <- rgrass7::execGRASS("g.list", type = "raster", pattern = "*inf*", mapset = ms) %>%
  attr("resOut")
rgrass7::use_sp()
# function to find the layer of interest
source("find_layers_GRASS.R")

# annotate
for (i in 1:nrow(layers_parms)) {
  # parms
  inf <- layers_parms$infrastructure[i]
  meas <- layers_parms$measure[i]
  shp <- layers_parms$shape[i]
  zoi <- layers_parms$zoi[i]

  # find layer name
  list_patterns <- list(paste0("^",inf), meas, shp, paste0(zoi,"$"))
  if(shp == "log") list_patterns <- list(paste0("^",inf), meas, shp)
  var_name <- find_layer_GRASS(list_patterns, 
                               layers_grass = layers)
  # read raster
  var <- rgrass7::readRAST(var_name, plugin=FALSE) %>%
    raster::raster() %>%
    terra::rast()
  # col name
  new_col_name <- paste(inf, meas, shp, zoi, sep = "_")
  if(shp == "log") new_col_name <- paste(inf, meas, shp, sep = "_")
  # extract
  dat[[new_col_name]] <- terra::extract(var, dat[,c("x33", "y33")], list = FALSE)[,2]
  # Ok!
  print(paste("OK:", new_col_name))
}

# save dataset
filename <- "cuminf_zoi_GPS_dataset_annotated.rda"
save(dat, file = paste0("../data/analysis_GPS/", filename))
```

Finally, we visualized the data within the study area.

```{r viz2, echo=TRUE, eval=FALSE}
dat_vect <- dat %>%
  dplyr::select(x33, y33) %>%
  dplyr::filter(use == 1) %>% 
  sf::st_as_sf(coords = c(1,2), crs = st_crs(study_area))

tm_shape(var) +
  tm_raster() +
  tm_shape(study_area_buff50) +
  tm_borders() +
  tm_shape(study_area) +
  tm_borders() +
  # tm_shape(dat_vect) +
  # tm_dots(size = 0.0001) +
  tm_scale_bar(breaks = c(15, 30, 45, 60))
```

