---
title: "Appendix B"
subtitle: "Applying the cumulative zone of influence approach to ecological studies"
author: ""
date: ""
geometry: margin=2cm
abstract: | 
  In this document, we provide examples of potential applications of the cumulative zone of
  influence approach in ecological studies with different contexts, biological measurements, and
  studied taxa. This is a supplementary information for Niebuhr et al. *Estimating the cumulative 
  impact and zone of influence of anthropogenic features on biodiversity*. Methods in Ecology
  and Evolution.
output:
  NinaR::jensAnalysis:
  highlight: tango
  fig_caption: yes
  toc: yes
header-includes:
  \usepackage{float}
  \usepackage{caption}
  \DeclareCaptionLabelFormat{Sformat}{#1 B#2}
  \captionsetup[figure]{labelformat=Sformat}
  \captionsetup[table]{labelformat=empty}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
library(NinaR)
library(kableExtra)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/",
  fig.align="center",
  fig.pos = 'H',
  fig.width = 5,
  fig.asp = 0.7,
  message=FALSE,
  warning=FALSE,
  options(scipen=999)
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())

# captions
library(captioner)
caption <- captioner("Table B", FALSE)

library(tibble)
```

# Introduction

In the main text, we present an application of the cumulative Zone of
Influence (ZOI) approach on habitat selection, using reindeer as a model
study. As mentioned there, the approach might be applied way beyond
animal movement ecology studies. In principle, the approach is useful in
any context in which predictor variables (sources of disturbance,
infrastructure, or any other spatial variable) vary in space and are
expected to affect the biological response variable or process being
studied beyond their actual limits. This includes but is not limited to
those studies in which either the distance to the nearest feature is
computed and used as a predictor (e.g. distance to roads in Torres et
al. 2016, or distance to forest edges in Martello et al. 2022) or those
in which the density or amount of a landscape variable is computed (e.g.
Jackson & Fahrig 2015, Miguet 2016, Martin 2018). Review literature
shows that the so-called multiscale approaches have been applied for a
different set of taxa (MacGarigal et al. 2016, Martin 2018) using
distinct biological measures for a wide range of ecological processes
(Martin 2016). Here we present examples of contexts in which our
approach might be useful to estimate cumulative impacts and zones of
influence, and provide studies to illustrate questions in which these
dimensions could have been evaluated.

# Illustrative examples

In table B1, we present some examples of how the cumulative impacts of
landscape disturbance and human infrastructure could be inferred on
different ecological processes, using distinct types of data, for
different response variables and organisms. For instance, studies on
movement and behavioral ecology may evaluate how movement rates (e.g.
Hansbauer et al. 2008) or the probability of an animal being in a given
behavioral state (e.g. Morales et al. 2004) depend on the influence of
disturbances in the landscape, and this influence might be measure
through either the ZOI of the nearest feature or the cumulative ZOI
metrics. Such response measures can based on different types of data
(e.g. locations recorded through GPS, VHF, or other devices,
capture-mark-recapture and camera trap data) for different organisms,
from animals who move actively (e.g. Ramos et al. 2020) to seeds, fungal
spores, or other propagules that are displaced by animals, wind, or
other agents (e.g. Carlo et al 2015, Norros et al. 2012).

Likewise, in the context of populations and communities, for example,
the ZOI metrics related to disturbance sources might be computed around
sampling points, considering multiple radii, and used to fit models
estimating their effects on population-level metrics (growth rate,
survival/mortality, fecundity; e.g. Cerqueira et al. 2021) for basically
any species or on community-level metrics (species diversity metrics,
species abundance, co-occurrence probability, interaction network
metrics; e.g. Torres et al. 2016).

More examples might be found in Table B1. We also provide in the table a
few references illustrating studies in which a similar approach was used
or could be adapted to incorporate an evaluation of cumulative impacts.
In most of these studies, though, only one of the metrics was used - the
distance to the nearest infrastructure or land cover type, related to
the ZOI of the nearest feature, or the density or amount of landscape
variables, related to the cumulative ZOI as presented here.

```{r table_applications, echo = F}
df <- tibble::tribble(
  ~`Ecological process`, ~`Response variable`, ~`Type of data`, ~`Statistical model`, ~`Taxon`, ~`Reference`,
  "Movement ecology", "Movement rate, step length, dispersal distance, turning angles", "GPS, VHF, Capture-mark- recapture", "Generalized linear (mixed) models [glm/glmm]", "Animals, seeds, fungi, other propagules", "[7, 8, 10, 11, 13]",
  "Habitat selection\nResource- or step- selection functions", "Presence, given availability", "GPS", "Binomial glm/glmm, conditional regression", "Animals", "[6, 14, 15, 16, 17]",
  "Ecological niche/\nSpecies distribution models", "Presence only, presence- background", "Ocurrence, camera-trap", "glm, MaxEnt, RandomForest", "Any organism", "[16, 18, 19, 20, 21]", 
  "Population dynamics", "Population size, growth rate, survival, mortality, fecundity", "Occurrence, count, distance-sampling, camera-trap", "glm/glmm, occupancy models, survival models", "Any organism", "[5, 12, 22]",
  "Community dynamics", "Species (co)occurrence, species occurrence probability, species richness, species abundances, diversity indices", "Multi-species occurrences or abundance", "Joint species distribution models", "Multiple", "[1, 2, 23, 24]",
  "Ecological interactions", "Interaction network metrics, probability of interaction", "Focal sampling of interactions", "glm/glmm", "Any type of interaction (e.g. throphic, mutualism, comensalism)", "[25, 26, 27, 28]",
  "Genetics", "Kinship, effective population size, genetic diversity, inbreeding coefficients, genetic differentiation between populations", "Genetic samples from multiple populations or areas", "Multiple", "Any organism", "[16, 29]",
  "Structural habitat, fragmentation, pollution", "Habitat edge influence measures, amount of pollution particles, bioaccumulation, concentration of pesticides in the water", "Multiple", "Multiple", NA, "[30, 31]"
)

kableExtra::kable(df, booktabs = T, caption = caption("Tab_1", "Examples of cases in which the cumulative impact approach could be applied, considering different ecological processes, response variables, types of data, statistical models, and organisms studied. The last column shows some references in which a similar approach was used or could have been applied.")) |> 
  kableExtra::kable_styling(full_width = T)
```

Looking more broadly, the approach can still be useful in other contexts
in which landscape variables measures at different scales might play a
role, e.g. in environmental research involving habitat structural and
abiotic measures (Lowe et al. 2022). For instance, the amount of
different pollutants in landscapes with different amounts of forest,
urban area, and roads could be evaluated, or the amount of pesticides
identified in the water could be tested against the proximity and
cumulative influence of agricutural and forested areas to the sampling
sites (e.g Liu and Yang 2018).

# Using the approach

To apply the cumulative ZOI approach in these different contexts, we
might start by following the same workflow presented in Fig. 2 of the
main text. First, using the spatial variables of interest, we should
compute the two ZOI metrics for multiple radii and shapes, as relevant
for the empirical case. Once this is done, they can be used to annotate
the biological data considering their spatial locations or sampling
sites. Given the table with response and predictor variable data is
ready, equation 1 from the main text might be adapted to look for the
relationship between the sampled biological response variables and the
ZOI and other spatial or non-spatial variables. Here the analyst might
use whatever modeling approach suits the data and context better; the
possibility of cumulative effects of each single spatial variable type
and their scale of effect or zone of influence are already considered
since they were used to compute the ZOI predictor variables. This means
that, in this formulation, it is not necessary to change the model
specification to assess the cumulative impacts of multiple features of a
given type of variable. More details on how this is done in practice are
shown in the [vignettes](https://ninanor.github.io/oneimpact/index.html)
of the `oneimpact` R package.

# References

[1] Torres, A., Jaeger, J. A. G., & Alonso, J. C. (2016). Assessing
large-scale wildlife responses to human infrastructure development.
Proceedings of the National Academy of Sciences, 113(30), 8472--8477.
\url{https://doi.org/10.1073/pnas.1522488113}

[2] Martello, F., Andriolli, F. S., Medeiros, H. R., Barão, T., &
Ribeiro, M. C. (2022). Edge contrast modulates ant community responses
to edge distance in agricultural landscapes. Agricultural and Forest
Entomology, 24(3), 289--300. \url{https://doi.org/10.1111/afe.12492}

[3] Jackson, H. B., & Fahrig, L. (2015). Are ecologists conducting
research at the optimal scale? Global Ecology and Biogeography, 24(1),
52--63. \url{https://doi.org/10.1111/geb.12233}

[4] Miguet, P., Jackson, H. B., Jackson, N. D., Martin, A. E., & Fahrig,
L. (2016). What determines the spatial extent of landscape effects on
species? Landscape Ecology, 31(6), 1177--1194.
\url{https://doi.org/10.1007/s10980-015-0314-1}

[5] Martin, A. E. (2018). The Spatial Scale of a Species' Response to
the Landscape Context Depends on which Biological Response You Measure.
Current Landscape Ecology Reports, 3(1), 23--33.\
\url{https://doi.org/10.1007/s40823-018-0030-z}

[6] McGarigal, K., Wan, H. Y., Zeller, K. A., Timm, B. C., & Cushman, S.
A. (2016). Multi-scale habitat selection modeling: A review and outlook.
Landscape Ecology, 31(6), 1161--1175.
\url{https://doi.org/10.1007/s10980-016-0374-x}

[7] Hansbauer, M. M., Storch, I., Leu, S., Nieto-Holguin, J.-P.,
Pimentel, R. G., Knauer, F., & Metzger, J. P. W. (2008). Movements of
neotropical understory passerines affected by anthropogenic forest edges
in the Brazilian Atlantic rainforest. Biological Conservation, 141(3),
782--791. \url{https://doi.org/10.1016/j.biocon.2008.01.002}

[8] Morales, J. M., Haydon, D. T., Frair, J., Holsinger, K. E., &
Fryxell, J. M. (2004). Extracting more out of relocation data: Building
movement models as mixtures of random walks. Ecology, 85(9), 2436--2445.
\url{https://doi.org/10.1890/03-0269}

[9] Ramos, D. L., Pizo, M. A., Ribeiro, M. C., Cruz, R. S., Morales, J.
M., & Ovaskainen, O. (2020). Forest and connectivity loss drive changes
in movement behavior of bird species. Ecography, 43(8), 1203--1214.
\url{https://doi.org/10.1111/ecog.04888}

[10] Carlo, T. A., Tewksbury, J. J., & del Río, C. M. (2009). A new
method to track seed dispersal and recruitment using 15 N isotope
enrichment. Ecology, 90(12), 3516--3525.
\url{https://doi.org/10.1890/08-1313.1}

[11] Norros, V., Penttilä, R., Suominen, M., & Ovaskainen, O. (2012).
Dispersal may limit the occurrence of specialist wood decay fungi
already at small spatial scales. Oikos, 121(6), 961--974.\
\url{https://doi.org/10.1111/j.1600-0706.2012.20052.x}

[12] Cerqueira, R. C., de Rivera, O. R., Jaeger, J. A. G., & Grilo, C.
(2021). Direct and indirect effects of roads on space use by jaguars in
Brazil. Scientific Reports, 11(1), 22617.
\url{https://doi.org/10.1038/s41598-021-01936-6}

[13] van Moorter, B., Bunnefeld, N., Panzacchi, M., Rolandsen, C. M.,
Solberg, E. J., & Sæther, B.-E. (2013). Understanding scales of
movement: Animals ride waves and ripples of environmental change.
Journal of Animal Ecology, 82(4), 770--780.
\url{https://doi.org/10.1111/1365-2656.12045}

[14] Laforge, M. P., Brook, R. K., van Beest, F. M., Bayne, E. M., &
McLoughlin, P. D. (2016). Grain-dependent functional responses in
habitat selection. Landscape Ecology, 31(4), 855--863.\
\url{https://doi.org/10.1007/s10980-015-0298-x}

[15] Zeller, K. A., Vickers, T. W., Ernest, H. B., & Boyce, W. M.
(2017). Multi-level, multi-scale resource selection functions and
resistance surfaces for conservation planning: Pumas as a case study.
PLOS ONE, 12(6), e0179570.
\url{https://doi.org/10.1371/journal.pone.0179570}

[16] Zeller, K. A., Jennings, M. K., Vickers, T. W., Ernest, H. B.,
Cushman, S. A., & Boyce, W. M. (2018). Are all data types and
connectivity models created equal? Validating common connectivity
approaches with dispersal data. Diversity and Distributions, 24(7),
868--879. \url{https://doi.org/10.1111/ddi.12742}

[17] Skarin, A., Sandström, P., & Alam, M. (2018). Out of sight of wind
turbines: Reindeer response to wind farms in operation. Ecology and
Evolution, 8(19), 9906--9919. \url{https://doi.org/10.1002/ece3.4476}

[18] Panzacchi, M., van Moorter, B., Strand, O., Loe, L. E., & Reimers,
E. (2015). Searching for the fundamental niche using individual-based
habitat selection modelling across populations. Ecography, 38(7),
659--669. \url{https://doi.org/10.1111/ecog.01075}

[19] Rezaei, S., Mohammadi, A., Malakoutikhah, S., & Khosravi, R.
(2022). Combining multiscale niche modeling, landscape connectivity, and
gap analysis to prioritize habitats for conservation of striped hyaena
(Hyaena hyaena). PLOS ONE, 17(2), e0260807.
\url{https://doi.org/10.1371/journal.pone.0260807}

[20] Bertassoni, A., Costa, R. T., Gouvea, J. A., Bianchi, R. de C.,
Ribeiro, J. W., Vancine, M. H., & Ribeiro, M. C. (2019). Land-use
changes and the expansion of biofuel crops threaten the giant anteater
in southeastern Brazil. Journal of Mammalogy, 100(2), 435--444.
\url{https://doi.org/10.1093/jmammal/gyz042}

[21] Santos, J. P., Sobral-Souza, T., Brown Jr, K. S., Vancine, M. H.,
Ribeiro, M. C., & Freitas, A. V. L. (2020). Effects of landscape
modification on species richness patterns of fruit-feeding butterflies
in Brazilian Atlantic Forest. Diversity and Distributions, 26(2),
196--208. \url{https://doi.org/10.1111/ddi.13007}

[22] Regolin, A. L., Oliveira-Santos, L. G., Ribeiro, M. C., & Bailey,
L. L. (2021). Habitat quality, not habitat amount, drives mammalian
habitat use in the Brazilian Pantanal. Landscape Ecology, 36(9),
2519--2533. \url{https://doi.org/10.1007/s10980-021-01280-0}

[23] Miguet, P., Fahrig, L., & Lavigne, C. (2017). How to quantify a
distance‐dependent landscape effect on a biological response. Methods in
Ecology and Evolution, 8(12), 1717--1724.
\url{https://doi.org/10.1111/2041-210X.12830}

[24] Suárez-Castro, A. F., Raymundo, M., Bimler, M., & Mayfield, M. M.
(2022). Using multi-scale spatially explicit frameworks to understand
the relationship between functional diversity and species richness.
Ecography, 2022(6), e05844. \url{https://doi.org/10.1111/ecog.05844}

[25] Moreira, E. F., Boscolo, D., & Viana, B. F. (2015). Spatial
Heterogeneity Regulates Plant-Pollinator Networks across Multiple
Landscape Scales. PLOS ONE, 10(4), e0123628.
\url{https://doi.org/10.1371/journal.pone.0123628}

[26] Schneiberg, I., Boscolo, D., Devoto, M., Marcilio-Silva, V.,
Dalmaso, C. A., Ribeiro, J. W., Ribeiro, M. C., Guaraldo, A. de C.,
Niebuhr, B. B., & Varassin, I. G. (2020). Urbanization homogenizes the
interactions of plant-frugivore bird networks. Urban Ecosystems, 23(3),
457--470. \url{https://doi.org/10.1007/s11252-020-00927-1}

[27] Monteiro, E. C. S., Pizo, M. A., Vancine, M. H., & Ribeiro, M. C.
(2021). Forest cover and connectivity have pervasive effects on the
maintenance of evolutionary distinct interactions in seed dispersal
networks. Oikos, oik.08240. \url{https://doi.org/10.1111/oik.08240}

[28] Marjakangas, E., Abrego, N., Grøtan, V., Lima, R. A. F., Bello, C.,
Bovendorp, R. S., Culot, L., Hasui, É., Lima, F., Muylaert, R. L.,
Niebuhr, B. B., Oliveira, A. A., Pereira, L. A., Prado, P. I., Stevens,
R. D., Vancine, M. H., Ribeiro, M. C., Galetti, M., & Ovaskainen, O.
(2020). Fragmented tropical forests lose mutualistic plant--animal
interactions. Diversity and Distributions, 26(2), 154--168.
\url{https://doi.org/10.1111/ddi.13010}

[29] Collevatti, R. G., dos Santos, J. S., Rosa, F. F., Amaral, T. S.,
Chaves, L. J., & Ribeiro, M. C. (2020). Multi-Scale Landscape Influences
on Genetic Diversity and Adaptive Traits in a Neotropical Savanna Tree.
Frontiers in Genetics, 11.
\url{https://www.frontiersin.org/articles/10.3389/fgene.2020.00259}

[30] Dodonov, P., Harper, K. A., & Silva-Matos, D. M. (2013). The role
of edge contrast and forest structure in edge influence: Vegetation and
microclimate at edges in the Brazilian cerrado. Plant Ecology, 214(11),
1345--1359.

[31] Liu, Z., & Yang, H. (2018). The impacts of spatiotemporal landscape
changes on water quality in Shenzhen, China. International Journal of
Environmental Research and Public Health, 15(5), 1038.
<https://doi.org/10.3390/ijerph15051038>

[32] Lowe, E. B., Iuliano, B., Gratton, C., & Ives, A. R. (2022).
'Scalescape': An R package for estimating distance-weighted landscape
effects on an environmental response. Landscape Ecology, 37(7),
1771--1785. <https://doi.org/10.1007/s10980-022-01437-5>
