---
title: "Appendix 4"
subtitle: "Cumulative influence of infrastructure on reindeer space use: calculating influence measures"
author: "Bernardo Niebuhr, Bram van Moorter, Manuela Panzacchi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=2cm
abstract: | 
  In this document, we prepare covariates and wild reindeer movement data for habitat selection 
  analysis presented in Niebuhr et al. *Estimating the cumulative impacts and the zone of influence 
  from multiple anthropogenic infrastructure on biodiversity.* The main goal of this document is to show
  how to calculate the cumulative influence and the influence from the nearest feature for several
  infrastructure variables using the R-GRASS interface and the `oneimpact` package.
output:
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
library(NinaR)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/",
  fig.align="center",
  message=FALSE,
  warning=FALSE
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())
```

\newpage


In this document we prepare covariates and wild reindeer movement data for the Hardangervidda area for habitat
selection analysis. The reindeer GPS dataset is organized in a setup of use-availability, where used locations (column 
`use = 1`) are compared to locations available but not used by the animals (column `use = 0`). 
We annotate the GPS dataset with the following covariates:

- Private cottages
- Large tourist cabins
- Small tourist cabins
- Roads of high traffic
- Roads of low traffic
- Trails
- Land cover
- 4 PCAs corresponding to bio-climatic-geographical variation in Norway

For the infrastructure covariates, we calculate the cumulative influence and the influence of the nearest feature 
considering different influence functions (threshold, linear decay, Gaussian decay, and exponential decay) and
different zones of influence, from 100 m to 20 km.

Below we present the steps for this data preparation.

\newpage

# Load packages

We start loading the packages necessary for the analysis. Spatial analyses are done through GRASS GIS using
the `oneimpact` package.

```{r load_packages, eval=FALSE, message=FALSE, warning=FALSE}
#Load packages
library(NinaR)

# data manipulation
library(dplyr)
library(purrr)
library(ggplot2) # for plot
library(ggpubr) # for plot

# spatial packages
library(rgrass7)
library(terra)
library(sf)
library(tmap)

# oneimpact package
library(oneimpact)
```

# Connect to GRASS

We start by connecting the R session to GRASS GIS. We link it to the GRASS GIS database from the Norwegian
Institute for Nature Research (NINA), where the spatial data is already organized.

```{r connect_grass, echo=TRUE, eval=FALSE}
# Connect to GRASS GIS location and mapset
ms <- "u_bb_cuminf"
NinaR::grassConnect(mapset = ms)

# set flags
flags <- c("quiet", "overwrite")
```

\newpage


# Set the study area

Here we subset the vector with the limits of the study area, the Hardangervidda wild reindeer
area, and make a 50 km buffer around it, for preparing the spatial data considering also the surrounding
of the study area and avoid edge effects in the spatial data preparation.

```{r load_vectors, echo=TRUE, eval=FALSE}
# load study area
areas_g <- "wild_reindeer_areas@u_bram.van.moorter"
study_area_n <- "Hardangervidda"

# load study area
study_area <- rgrass7::read_VECT(areas_g) %>%
  sf::st_as_sf() %>% 
  dplyr::filter(name_area == study_area_n) %>%
  sf::st_union()
# plot(study_area)

# buffer around the study area
buff <- 50e3 # 50km buffer
study_area_buff50 <- study_area %>%
  sf::st_buffer(dist = buff)
# plot(study_area_buff50)

# put study area and buffer into the GRASS GIS mapset
study_area_g <- "study_area"
rgrass7::writeVECT(study_area, study_area_g,
                   v.in.ogr_flags = flags)
study_area_buff_g <- "study_area_buff50"
rgrass7::writeVECT(study_area_buff50, study_area_buff_g,
                   v.in.ogr_flags = flags)
```

Just a plot of the study area.

```{r viz1, echo=TRUE, eval=FALSE}
tm_shape(study_area_buff50) +
  tm_borders() +
  tm_shape(study_area) +
  tm_borders() +
  tm_scale_bar(breaks = c(15, 30, 45, 60))
```
\newpage

# Rasterize infrastructure vectors

Now we rasterize the vector layers corresponding to the infrastructure variables we'll
consider for this area. We use a previously created raster file with 100 m resolution as
a base grid for rasterizing the infrastructure layers.

For the private cottage and tourist cabins, we create rasters with value 1 where there
are infrastructure and NA where they are absent, to be used as input for the calculation
of the influence of the nearest feature. Here it is not relevant to count the number of features,
only to know if they are present or not in a given pixel, since this measure takes into
account of the influence of the nearest feature.

We also create rasters counting the number of cottages in each pixel, which are used as 
input for the calculation of the cumulative influence. This procedure is made with the 
`util_v2rast_count_GRASS` function of the `oneimpact` package.

```{r rasterize_maps_count, echo=TRUE, eval=FALSE}
# set region as the study area buffer
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# vectors to be rasterized
vectors_to_rasterize_c <- c("private_cabins@p_prodchange_envpoints",
                            "pub_cabins_summer_high@p_prodchange_envpoints",
                            "pub_cabins_summer_low@p_prodchange_envpoints")
# output raster map names
outputs_rasterize_c1 <- c("private_cabins",
                          "public_cabins_high",
                          "public_cabins_low")

# option
use_rasterize_c1 <- "val"

#---
# First rasterize using val=1 to use as input for calc_influence_nearest

# put all in a tibble
rasterize_parms_c1 <- tibble::tibble(input = vectors_to_rasterize_c,
                                     output = outputs_rasterize_c1,
                                     use = use_rasterize_c1)

# rasterize
rasterized_c1 <- rasterize_parms_c1 %>% purrr::pmap(rgrass7::execGRASS, cmd = "v.to.rast",
                                                    flags = flags)

# colors
f <- tempfile()
cat("1 grey", file = f)
# readr::read_file(f)
colored <- purrr::map(outputs_rasterize_c1, function(x)
  rgrass7::execGRASS(cmd = "r.colors", map = x, rules = f, flags = ))

#---
# Then rasterize using count to use as input for calc_influence_cumulative

# output raster map names
outputs_rasterize_c2 <- c("private_cabins_count",
                          "public_cabins_high_count",
                          "public_cabins_low_count")

# parameters
count_parms <- tibble::tibble(x = vectors_to_rasterize_c,
                              output = outputs_rasterize_c2,
                              set_region = FALSE)

# rasterize with count
count <- count_parms %>% purrr::pmap(oneimpact::util_v2rast_count_GRASS,
                                     quiet = F, overwrite = TRUE)

```

For the roads and trails, however, for 100m pixels there can hardly be more than one 
feature in each pixel, and if this occurs it is not relevant for the analysis. Therefore, 
a rasterization counting the number of features is not made. Instead, besides the common
rasterization used to create the inputs for the calculation of the influence of
the nearest feature, we create binary raster maps (1 where there are infrastructure, 0 
elsewhere) to be used as input for the cumulative influence calculation.

```{r rasterize_maps_value1, echo=TRUE, eval=FALSE}
# set region as the study area buffer
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# vectors to be rasterized
vectors_to_rasterize_1 <- c("roads_summer_high@p_prodchange_roads",
                            "roads_summer_low@p_prodchange_roads",
                            "trails@p_RenRein_trails2")
# output raster map names
outputs_rasterize_1 <- c("roads_high",
                         "roads_low",
                         "trails")

# option - values = 1
use_rasterize_1 <- c("val")

# put all in a tibble
rasterize_parms_1 <- tibble::tibble(input = vectors_to_rasterize_1,
                                    output = outputs_rasterize_1,
                                    use = use_rasterize_1)

# rasterize
rasterized_1 <- rasterize_parms_1 %>% purrr::pmap(rgrass7::execGRASS, cmd = "v.to.rast",
                                                  flags = flags)

# colors
f <- tempfile()
cat("1 grey", file = f)
# readr::read_file(f)
colored <- purrr::map(outputs_rasterize_1, function(x)
  rgrass7::execGRASS(cmd = "r.colors", map = x, rules = f, flags = "quiet"))

# binarize maps for cumulative influence calculation
output_bin_1 <- paste0(outputs_rasterize_1, "_bin")

binarize_parms_1 <- tibble::tibble(x = outputs_rasterize_1,
                                   output = output_bin_1,
                                   null = 0)

binarized_1 <- binarize_parms_1 %>% purrr::pmap(oneimpact::util_binarize_GRASS,
                                                quiet = TRUE, overwrite = TRUE)
binarized_1 <- unlist(binarized_1)
```

```{r cut_and_save_layers, echo=FALSE, eval=FALSE}
vect_export <- c(vectors_to_rasterize_c, vectors_to_rasterize_1)
vect_export_within <- strsplit(vect_export, "@") %>% 
  sapply(first) %>% 
  paste(., "within", sep = "_")
study_area_g <- "study_area"
out_names <- gsub("within", "v.gpkg", vect_export_within)

# cut and retrieve maps for the study area
layers_in_r <- list()
for(i in 2:length(vect_export)) {
  cut
  rgrass7::execGRASS("v.select", ainput = vect_export[i], binput = study_area_g,
                     output = vect_export_within[i], operator = "within", flags = "overwrite")
  retrieve to R
  layers_in_r[[i]] <- rgrass7::read_VECT(vect_export_within[i])
  # export from GRASS
  # rgrass7::execGRASS("v.out.ogr", input = vect_export_within[i], 
  #                    output = paste0("../data/analysis_GPS/", out_names[i]))
}

# export from R
for(i in 2:length(out_names)) {
  terra::writeVector(layers_in_r[[i]], filename = paste0("../data/analysis_GPS/", out_names[i]), 
                     overwrite = TRUE)
}
```

\newpage


# Calculate the influence of the nearest feature

Here we calculate, for each infrastructure type, the influence of the nearest feature,
for different influence functions and zones of influence. To do so, we use the function
`calc_influence_nearest` from the `oneimpact` package. We also calculate the log-distance
from the nearest feature, which is commonly used in ecological studies, to compare its
performance in model fitting with the other influence variables.

```{r calculate_nearest, echo=TRUE, eval=FALSE}
# calculate influence from the nearest feature
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000, 30000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# define parameters
nearinf_parms <- expand.grid(zoi = zoi_vals, type = shapes)

# calculate influence from the nearest feature

# private cabins
nearinf_private_cabins <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                        x = "private_cabins",
                                                        where = "GRASS",
                                                        quiet = TRUE, overwrite = TRUE)

# public cabins high
nearinf_public_cabins_high <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                            x = "public_cabins_high",
                                                            where = "GRASS",
                                                            quiet = TRUE, overwrite = TRUE)

# public cabins low
nearinf_public_cabins_low <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                           x = "public_cabins_low",
                                                           where = "GRASS",
                                                           quiet = TRUE, overwrite = TRUE)

# roads high
nearinf_roads_high <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                    x = "roads_high",
                                                    where = "GRASS",
                                                    quiet = TRUE, overwrite = TRUE)

# roads low
nearinf_roads_low <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                   x = "roads_low",
                                                   where = "GRASS",
                                                   quiet = TRUE, overwrite = TRUE)

# summer trails
nearinf_trails <- nearinf_parms %>% purrr::pmap(oneimpact::calc_influence_nearest,
                                                x = "trails",
                                                where = "GRASS",
                                                quiet = TRUE, overwrite = TRUE)

# additionally, also calculate log-distance
infrastructure <- c("private_cabins", "public_cabins_high", "public_cabins_low", 
                    "roads_high", "roads_low", "trails")

# calculate log-distance from the nearest feature for all infrastructure
nearinf_log <- purrr::map(infrastructure, oneimpact::calc_influence_nearest,
                          type = "log", where = "GRASS",
                          quiet = TRUE, overwrite = TRUE)
```

\newpage


# Calculate the cumulative influence

Here we calculate, for each infrastructure type, the cumulative influence of multiple features,
for different influence functions and zones of influence. To do so, we use the function
`calc_influence_cumulative` from the `oneimpact` package.

```{r calculate_cumulative, echo=TRUE, eval=FALSE}
# calculate cumulative influence

# input maps, ZoI values, and filter shapes
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000, 30000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# define parameters
cuminf_parms <- expand.grid(zoi = zoi_vals, type = shapes)

# calculate cumulative influence

# private cabins
cuminf_private_cabins <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                                      x = "private_cabins_count",
                                                      where = "GRASS",
                                                      quiet = TRUE, overwrite = TRUE)

# public cabins high
cuminf_public_cabins_high <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                                          x = "public_cabins_high_count",
                                                          where = "GRASS",
                                                          quiet = TRUE, overwrite = TRUE)

# public cabins low
cuminf_public_cabins_low <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                                         x = "public_cabins_low_count",
                                                         where = "GRASS",
                                                         quiet = TRUE, overwrite = TRUE)

# roads high
cuminf_roads_high <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                                  x = "roads_high_bin",
                                                  where = "GRASS",
                                                  quiet = TRUE, overwrite = TRUE)

# roads low
cuminf_roads_low <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                                 x = "roads_low_bin",
                                                 where = "GRASS",
                                                 quiet = TRUE, overwrite = TRUE)

# summer trails
cuminf_trails <- cuminf_parms %>% purrr::pmap(oneimpact::calc_influence_cumulative,
                                              x = "trails_bin",
                                              where = "GRASS",
                                              quiet = TRUE, overwrite = TRUE)

```

\newpage


# Prepare and annotating GPS data

Finally, we load the reindeer GPS data, prepare it fr analysis, and annotate it with
the infrastructure data to allow assessing their effects on reindeer space use.

The reindeer data was already organized in a use-availability design, where used locations (column 
`use = 1`) are compared to locations available but not used by the animals (column `use = 0`). It 
was also already annotated with land cover classes (that were recombined here) and the PCA climatic
variables.

```{r read_gps, echo=TRUE, eval=FALSE}
# load gps data
filename <- "wrein_dat_sub_20200518V2.rda"
load(file=paste0("/data/R/Prosjekter/Rein/RSF_output/", filename)) #servers, e.g. ninrstudio03
study_area_n <- "Hardangervidda"

# subset summer
dat <- dat$sum
names(dat)
dat <- dat[dat$herd == study_area_n,] 

# reclass land cover
dat$NORUTreclass <- ifelse(dat$norut<9, "11forest", ifelse(dat$norut<12, "bog", 
                                                           ifelse(dat$norut<21, "mountain", 
                                                                  ifelse(dat$norut<22, "glacier", 
                                                                         ifelse(dat$norut<23, "water", 
                                                                                ifelse(dat$norut<25,"other", NA))))))
dat$NORUTreclass <- ifelse(dat$NORUTreclass=="mountain", dat$norut, dat$NORUTreclass)
tapply(dat$use, dat$NORUTreclass, mean) # explore the proportion of each used point in classes
tapply(dat$use, dat$NORUTreclass, length) # explore the number of points in each class
sum(is.na(dat$NORUTreclass))

# remove old infrastructure columns
to_remove <- c("houses", "powerlines", "private_cab", "pub_cab", 
               "railway", "roads", "ski", "trails") %>%
  paste(collapse = "|")
dat <- dat %>%
  dplyr::select(-c(grep(to_remove, colnames(dat))))
names(dat)
```

Finally, we annotate the GPS dataset with the influence measures calculated in this 
document and save it externally for the statistical analysis.

```{r annotate_gps, echo=TRUE, eval=FALSE}
# variables
infrastructure = c("private_cabins", "public_cabins_high", "public_cabins_low", 
                   "roads_high", "roads_low", "trails")
measures <- c("cumulative", "nearest")
zoi_vals <- c(100, 250, 500, 1000, 2500, 5000, 10000, 20000)#, 30000)
shapes <- c("threshold", "bartlett", "exp_decay", "Gauss")

# parameters for the layers
layers_parms <- expand.grid(infrastructure = infrastructure,
                            measure = measures,
                            shape = shapes,
                            zoi = zoi_vals) %>%
  dplyr::arrange(infrastructure, measure, shape, zoi)
# add log distance to parameters
layers_parms <- layers_parms %>% 
  dplyr::bind_rows(
    expand.grid(infrastructure = infrastructure,
                measure = "nearest",
                shape = "log",
                zoi = NA))

# re-set region
rgrass7::execGRASS("g.region", vector = study_area_buff_g,
                   align = "houses_10000@p_prodchange_envpoints",
                   flags = "p")

# list all relevant layers
layers <- rgrass7::execGRASS("g.list", type = "raster", pattern = "*inf*", mapset = ms) %>%
  attr("resOut")
rgrass7::use_sp()
# function to find the layer of interest
source("find_layers_GRASS.R")

# annotate
for (i in 1:nrow(layers_parms)) {
  # parms
  inf <- layers_parms$infrastructure[i]
  meas <- layers_parms$measure[i]
  shp <- layers_parms$shape[i]
  zoi <- layers_parms$zoi[i]

  # find layer name
  list_patterns <- list(paste0("^",inf), meas, shp, paste0(zoi,"$"))
  if(shp == "log") list_patterns <- list(paste0("^",inf), meas, shp)
  var_name <- find_layer_GRASS(list_patterns, 
                               layers_grass = layers)
  # read raster
  var <- rgrass7::readRAST(var_name, plugin=FALSE) %>%
    raster::raster() %>%
    terra::rast()
  # col name
  new_col_name <- paste(inf, meas, shp, zoi, sep = "_")
  if(shp == "log") new_col_name <- paste(inf, meas, shp, sep = "_")
  # extract
  dat[[new_col_name]] <- terra::extract(var, dat[,c("x33", "y33")], list = FALSE)[,2]
  # Ok!
  print(paste("OK:", new_col_name))
}

# save dataset
filename <- "cuminf_zoi_GPS_dataset_annotated.rda"
save(dat, file = paste0("../data/analysis_GPS/", filename))
```

Finally, we visualize the data within the study area.

```{r viz2, echo=TRUE, eval=FALSE}
dat_vect <- dat %>%
  dplyr::select(x33, y33) %>%
  dplyr::filter(use == 1) %>% 
  sf::st_as_sf(coords = c(1,2), crs = st_crs(study_area))

tm_shape(var) +
  tm_raster() +
  tm_shape(study_area_buff50) +
  tm_borders() +
  tm_shape(study_area) +
  tm_borders() +
  # tm_shape(dat_vect) +
  # tm_dots(size = 0.0001) +
  tm_scale_bar(breaks = c(15, 30, 45, 60))
```

