---
title: "Simulate responses"
author: "Bernardo Niebuhr"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load_packages}
#Load packages
library(mobsim)
library(raster)
library(ZOItools)
library(sf)
library(boot)
```


## Simulate a landscape

First we simulate a landscape, using point-type infrastructure as an example.

```{r simulate_landscapes}
set.seed(1234)

# multiple focii
ext <- 30000
wd <- 0.1*ext
scales <- c(250, 500, 1000, 2500, 5000)/2
rasts <- simulate_dist_dens(n_features = 1000, centers = 5,
                            width = wd, res = 100,
                            extent_x = c(0, ext), extent_y = c(0, ext),
                            buffer_around = 10000,
                            type_density = "Gauss", scale = scales,
                            extent_x_cut = c(0, ext), extent_y_cut = c(0, ext))
plot(rasts)
```

## Create a response

Now we wish to simulate a response using this landscape. We'll consider a logistic response in the format

\begin{align*}
y_i \sim {\sf Binom}(p_i) \\
logit(p_i) = log \left( \frac{p_1}{1 - p_i} \right) = \beta_{0i} + \beta_1 \cdot x_{1i} + \beta_2 \cdot x_{2i}
\end{align*}

where $y_i$ is the response variable (1 = used and 0 = available in the context of habitat suitability modeling), 
$x_{1i}$ is a first variable independent from the space (added here just to consider one other factor of influence on
the output), $x_{2i}$ is either the `density` (the density of features) or `log(distance)` (the distance to the nearest feature), 
and $\beta_0$, $\beta_1$, and $\beta_2$ are the coefficients of the equation.

We'll simulate two scenarios. In the first one we set `log(distance)` as the variable generating the data and in the second one we set `density` as the variable generating the data. In each case, we then fit models considering each variable and we compare and interpret the results.

```{r}
n <- 500
beta <- c(2, 1, .5)
x1 <- runif(n)

coords <- rbind(c(0,0),c(ext,0),c(ext,ext),c(0,ext),c(0,0))
pts <- st_sf(st_sample(st_sfc(st_polygon(list(coords))), n))

vars <- raster::extract(rasts, pts)
vars[, "distance"] <- log(vars[, "distance"])
head(vars)

# simulate
x2 <- vars[,1] # dist
logit_p = beta[1] + (beta[2]*x1) + (beta[3]*x2)
y <- rbinom(n, 1, inv.logit(logit_p))

dat <- data.frame(x1, x2, y)
```

