---
title: "Multiple ways of simulating spatial point patterns"
author: "Bernardo Niebuhr"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we document multiple ways of simulating spatial point patterns and how to transform them into rasters.

## From the scratch

This code comes from the class notes from the course "Modelagem Estatistica para Ecologia e Recursos Naturais",
developed by the wonderful professors Paulo Inacio Prado and Joao Batista, from Sao Paulo University in Brazil.
Link [here](http://cmq.esalq.usp.br/BIE5781/doku.php?id=01-discretas:01-discretas).

```{r}
# Geracao de 20 centros de patches
set.seed(42)
cx <- runif(20, 0, 20)
cy <- runif(20, 0, 20)

# Distribuicao de 2000 pontos ao redor dos centros dos agregados, seguindo uma dist. normal
# padronizada (mu=0, sigma=1) ao redor deles
px <- rnorm(2000)
py <- rnorm(2000)
x1 <- cx+px
y1 <- cy+py
# c(1,2) + c(2,3, 4, 5) # Interessate! da pra somar uma serie de valores com outra com um 
# numero de elementos multiplo do numero de elementos da outra serie! (eles se alternam na soma)

# queremos so os pontos que estao dentro da area x,y = [0,20]
x2 <- x1[x1>0 & x1<20 & y1>0 & y1<20]
y2 <- y1[x1>0 & x1<20 & y1>0 & y1<20]

# Para reproduzir a distribuicao espacial das plantas
plot(x=x2,y=y2,axes=F, xlab="", ylab="", xlim=c(-1,20),ylim=c(0,21))
points(cx,cy, col="blue", pch=19)
segments(x0=rep(0,20),y0=0:20,x1=rep(20,20),y1=0:20, col="gray", lty=2)
segments(x0=0:20,y0=rep(0,20),x1=0:20,y1=rep(20,20),col="gray", lty=2)
text(y=seq(0.5,19.5,by=1),x=-1, labels=20:1)
text(x=seq(0.5,19.5,by=1),y=21, labels=LETTERS[1:20])

```


## Using the `mobsim` package

The package is originally intended to create these pattern to study multiple species in species abundance distribution 
studies. But we can set only 1 species and we get the intended results. This code is based on the examples and vignettes of the
`mobsim` package.

```{r mobsim}
# install.packages("mobsim", dep = T)
library(mobsim)

# random
comm_rand <- sim_poisson_community(s_pool = 1, n_sim = 300)
plot(comm_rand)

# agreggated, multiple focii (eg. cities)
comm_agg <- sim_thomas_community(s_pool = 1, n_sim = 100, 
                                 sigma = .05, mother_points = 4,
                                 xrange = c(0,1), yrange = c(0,1))
plot(comm_agg)

# can change the number of mother points

# agreggated, one focii (eg. wind park)
comm_agg <- sim_thomas_community(s_pool = 1, n_sim = 100, 
                                 sigma = .05, mother_points = 1,
                                 xrange = c(0,1), yrange = c(0,1))
plot(comm_agg)

# gradient
comm_agg <- sim_thomas_community(s_pool = 1, n_sim = 100, 
                                 sigma = .3, mother_points = 1,
                                 xrange = c(0,1), yrange = c(0,1))
plot(comm_agg)

# random
comm_agg <- sim_thomas_community(s_pool = 1, n_sim = 100, 
                                 sigma = .99, mother_points = 1,
                                 xrange = c(0,1), yrange = c(0,1))
plot(comm_agg)

```

## Using the `raster` package

Here the exercise is not to create the random points, but to rasterize them. The same may be made with the `terra` package, 
even though the function to `rasterize` points seem to be a little different.
This code is based on the examples from the `rasterize` function from the `raster` package, which may be found [here](https://rdrr.io/cran/raster/man/rasterize.html).

```{r raster}
library(raster)

# create random points
set.seed(123)
n <- 1000
x <- runif(n) * 360 - 180
y <- runif(n) * 180 - 90
xy <- cbind(x, y)
plot(xy)

# create raster
r <- raster(ncols=36, nrows=18)

# get the (last) indices
r0 <- rasterize(xy, r)
plot(r0)
# presence/absensce (NA) (is there a point or not?)
r1 <- rasterize(xy, r, field=1)
plot(r1)
# how many points?
r2 <- rasterize(xy, r, fun=function(x,...)length(x))
plot(r2)

vals <- runif(n)
# sum of the values associated with the points
r3 <- rasterize(xy, r, vals, fun=sum)
```

